On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-09-28 paper.tex
Moved section on another presentation to the preliminaries

\subsection{Relative Steinberg groups} \label{sec:relative-steinberg}
First of all, let us recall the definition of the category of pairs. 
As suggested by the notation, the objects of $\catname{Pairs}$ are pairs $(R, I)$ consisting of a unital commutative ring $R$ and an ideal $I \trianglelefteq R$.
The set of morphisms between $(R, I)$ and $(S, J)$ consists, by definition, of all ring-theoretic maps $f\colon R\to S$ such that $f(I)\subseteq J$.
There is an obvious fully faithful embedding $\iota\colon \catname{CRings} \to \catname{Pairs}$ defined on objects by $\iota(R) = (R, R)$.
%Now if there is a functor $F\colon \catname{CRings}\to \catname{Grp}$ then its {\it relativization} is another functor $\widetilde{F}\colon \catname{Pairs} \to \catname{Groups}$ such that $\widetilde{F} \iota = F$.

@@

%Now suppose that $t, u \in R^*$ are invertible elements such that $u-1 \in I$. 
In the sequel we will also need a {\it relative Steinberg symbol}, i.\,e. a map
\begin{equation*} \{-, -\}'_\alpha \colon R^* \times (1 + I)^* \to \St(\Phi, R, I) \end{equation*}
such that $ \ext{I} \{u, v\}'_\alpha = \{u, v\}_\alpha$. It is not hard to verify that the image under $\mu_I$ of the element
\begin{equation*} f(\alpha, t, u)= y_\alpha(t(u-1)) \cdot y_{-\alpha}(-t^{-1}(u^{-1}-1))^{x_\alpha(-t)} \cdot y_{\alpha}(t(u-1))^{w_\alpha(-t)}, \end{equation*}
equals $w_\alpha(tu) \cdot w_\alpha(-t)$. This allows us to set $\{t, u \}'_\alpha := f(\alpha, t, u) f(\alpha, 1, u)^{-1}.$
Now, since $\mu_I\{ t, u \}'_\alpha = w_\alpha(tu) w_\alpha(-t) w_\alpha(1) w_\alpha(-u) = \{t, u\}_\alpha,$ 
 we have defined the desired symbol.
\begin{rem}
 Our relative Steinberg symbols differs from the one defined in~\cite{Ste73} in that
  it takes values in $\St(\Phi, R, I)$ rather than the group $\St(\Phi, R)$. %Elaborate on this...
\end{rem}

@@

To simplify notation, throughout this section we set $N_1 = N(\varpi_1)$, $N^1 = N(-\varpi_1)$.
We only construct one of the maps from the definition of $\sigma$-pair, namely $\sigma(\varpi_1^\vee) \colon N_1 \to N^1$,
 for shortness we denote it by $\sigma_1$.

In order to define the desired map $\sigma_1$ we will use the alternate description of the subgroup $N_1 = N(\varpi_1^\vee)$
given in terms of ``another presentation''. We start with the following simple observation.
\begin{lemma} \label{lem:n1-decomp} For $n\geq 4$ there is an isomorphism $N_1 \cong N_{0} \rtimes P_1^-(A)$, 
 where $N_{0}$ denotes the subgroup $\St(n, A[t], tA[t])$ and $P_1^-(A)$ is the subgroup of $\St(n, A)$ generated by $x_{ij}(\xi)$ with $i\neq 1$.
\end{lemma}

Recall that the universal property of semidirect products asserts that for any group $H$ acting on a group $N$ via $\phi \colon H \to \Aut(N)$ 
and any group homomorphisms $f_N\colon N \to G$, $f_H\colon H \to G$ satisfying 
\begin{equation} \label{eq:coherence-condition} f_N(\phi(h)(n)) = f_H(h) f_N(n) f_H(h)^{-1},\ n\in N,\ h\in H,\end{equation} 
there exists a unique map $f\colon N \rtimes H \to G$ extending $f_N$ and $f_H$.

Thus, by the above lemma, in order to construct the map $\sigma_1$ we need to construct the maps 
\[ (\sigma_1)_{P_1^-(A)} \colon P_1^-(A) \to N^1, \ \ (\sigma_1)_{N_{0}} \colon N_{0} \to N^1\]
and then verify that they satisfy~\eqref{eq:coherence-condition}.

It is easy to define the first map $(\sigma_1)_{P_1^-(A)}$, indeed, using the decomposition $P_1^-(A) \cong U^-_1 \rtimes L_1$ where 
\[U^-_1 = \langle x_{i1}(\xi) \mid i\neq 1,\ \xi\in A \rangle \text{ and } L_1 = \langle x_{ij}(\xi) \mid i,  j \neq 1,\ \xi\in A\rangle \]
we apply the universal property of semidirect products once again 
and define $(\sigma_1)_{P_1^-(A)}$ by requiring that it acts identically on $L_1$ (notice that $L_1 \subseteq N^1$)
and acts on elements of $U^-_1$ via the formula $(\sigma_1)_{P_1^-(A)}(x_{i1}(\xi))= x_{i1}(t\xi)$.

Before we proceed with the construction of the second map $(\sigma_1)_{N_{1,0}}$ we need to recall several technical
 constructions pertaining to ``another presentation'' of the group $N_{0}$.

For $u \in R^n$ we denote by $D(u)$ the subset of $R^n$ consisting of all $v$ having at least two zero entries and satisfying $u^tv = 0$.
Recall from 3.2 of~\cite{Ka77} that for every $u, v, w \in R[t]^n$ such that $u^t \cdot v = 0$ there
 is a decomposition of $(w^t u) \cdot v$ into a sum of elements of $D(u)$, called {\it canonical decomposition}:
\setcounter{equation}{1}
\renewcommand{\theequation}{\arabic{equation}}
\begin{equation} \label{eq:canonical} (w^tu) \cdot v=\sum_{i<j}u_{ij} c_{ij}(v, w),\text{ where $u_{ij}=e_iu_j-e_ju_i,$ $c_{ij}(v, w)=v_iw_j-v_jw_i.$}\end{equation}

Let $A$ be a commutative ring and $I$ be a splitting ideal of $A$.
Let $u, v \in A^n$ be such that $u^t v = 0$ and assume, moreover, that either $u$ or $v$ has at least one zero entry.
Recall from 3.8, 3.10 of~\cite{Ka77} that under these assumptions one can define certain element
 $x(u, v)$ of $\St(n, A)$ such that $\pi(x(u, v)) = t(u, v) = 1 + uv^t$.
The following properties of $x(u, v)$ are standard (cf.~\cite[Lemma~1.1]{Tu83}).
\begin{lemma} \label{lem:xsmall-properties}
\begin{lemlist}
 \item \label{item:xsmall-scalar} If $v$ or $w$ has at least two zero components, then 
 \begin{equation}\nonumber x(v, wa) = x(va, w).\end{equation}
 \item \label{item:xsmall-additivity} If $w_1$ and $w_2$ have at least one common zero coordinate then
 \begin{equation}\nonumber x(v, w_1)x(v, w_2) = x(v, w_1+w_2).\end{equation}
 \item \label{item:xsmall-commute} If $v$, $v'$ are simultaneously orthogonal to $w$ and $w'$ and the elements $w, w'$
  both have at least two zero components then $x(v, w)$ and $x(v', w')$ commute.
\end{lemlist}
\end{lemma}

Let $u \in E(n, A)e_1$, $v \in I^n$ be such that $u^tv = 0$.
Recall from \S~4 of~\cite{LS17} that one can define the following elements of $\St(n, A, I)$:
\begin{equation} \label{eq:sigma-definition}
 F(u, v) = \prod\limits_{i=1}^r x(u,  v_i),\ \
 S(v, u) = \prod\limits_{i=1}^r x( v_i, u).
\end{equation}
Here $\{v_r\}$, $r=1,\ldots,N$ is any collection of elements of $I^n \cap D(u)$ such that $\sum_{r=1}^N v_r = v$.
Since $u$ is unimodular, such a collection always exists by~\eqref{eq:canonical}, moreover
 $F(u, v)$ and $S(u, v)$ do not depend on its choice. %TODO: Give reference

Recall from~\cite{LS17} that it is possible to give a presentation of the relative Steinberg group $\St(n, A, I)$ formulated in terms of elements $F(u, v)$, $S(v, u)$.
More formally, \cite[Proposition 3.10]{LS17} asserts that $\St(n,\,A,\,I)$ admits presentation with the set of generators
\begin{multline*} \{F(u,\,v)\mid u\in E(n,\,A)e_1,\ v\in I^n,\ u^tv=0\}\, \cup \\ \cup \{S(u,\,v)\mid u\in I^n,\ v\in E(n,\,A)e_1,\ u^tv=0\}\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}

\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)^{-1}=F(t(u,\,v)u',\,t(v,\,u)^{-1} v'), \label{conj3} \\
&F(ge_1,\,g^*e_2a)=S(ge_1a,\,g^*e_2),\ \text{for all}\ a\in I,\, g \in E(n, A).
\end{align}
Here $g^*$ denotes the contragradient matrix, i.\,e. $g^* = {g^{-1}}^t$ and $e_i$ is the standard basis vector of $A^n$.

\setcounter{equation}{3}
\renewcommand{\theequation}{\arabic{equation}}

To define the map $(\sigma_1)_{N_{1,0}}$ we will invoke the above presentation in the special case $A=R[t]$, $I=tR[t]$.
Indeed, define $\sigma_1$ on $N_{1,0}=\St(n, R[t], tR[t])$ by the following identities: 
\begin{equation}
 \sigma_1(F(u, v)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot t u, \delta_1 \cdot t^{-1} v_i),\ \
 \sigma_1(S(v, u)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot v_i, \delta_1 \cdot u).
\end{equation}

>>>>

\subsection{Relative Steinberg groups} \label{sec:relative-steinberg}
First of all, let us recall the definition of the category of pairs. 
As suggested by the notation, the objects of $\catname{Pairs}$ are pairs $(R, I)$ consisting of a unital commutative ring $R$ and an ideal $I \trianglelefteq R$.
The set of morphisms between $(R, I)$ and $(S, J)$ consists, by definition, of all ring-theoretic maps $f\colon R\to S$ such that $f(I)\subseteq J$.
There is an obvious fully faithful embedding $\iota\colon \catname{CRings} \to \catname{Pairs}$ defined on objects by $\iota(R) = (R, R)$.

@@

In the sequel we will also need a {\it relative Steinberg symbol}, i.\,e. a map
\begin{equation*} \{-, -\}'_\alpha \colon R^* \times (1 + I)^* \to \St(\Phi, R, I) \end{equation*}
such that $ \ext{I} \{u, v\}'_\alpha = \{u, v\}_\alpha$. It is not hard to verify that the image under $\mu_I$ of the element
\begin{equation*} f(\alpha, t, u)= y_\alpha(t(u-1)) \cdot y_{-\alpha}(-t^{-1}(u^{-1}-1))^{x_\alpha(-t)} \cdot y_{\alpha}(t(u-1))^{w_\alpha(-t)}, \end{equation*}
equals $w_\alpha(tu) \cdot w_\alpha(-t)$. This allows us to set $\{t, u \}'_\alpha := f(\alpha, t, u) f(\alpha, 1, u)^{-1}.$
Now, since $\mu_I\{ t, u \}'_\alpha = w_\alpha(tu) w_\alpha(-t) w_\alpha(1) w_\alpha(-u) = \{t, u\}_\alpha,$ 
 we have defined the desired symbol.
\begin{rem}
 Our relative Steinberg symbols differs from the one defined in~\cite{Ste73} in that
  it takes values in $\St(\Phi, R, I)$ rather than in $\St(\Phi, R)$. %Elaborate on this...
\end{rem}

\subsection{Another presentation of linear Steinberg groups} \label{sec:another-presentation}
%TODO: Formulate why do we write what we write
Throughout this subsection $R$ denotes a commutative ring and $I$ its splitting ideal.

For $u \in R^n$ we denote by $D(u)$ the subset of $R^n$ consisting of all $v$ having at least two zero entries and satisfying $u^tv = 0$.
Recall from 3.2 of~\cite{Ka77} that for every $u, v, w \in R[t]^n$ such that $u^t \cdot v = 0$ there
 is a decomposition of $(w^t u) \cdot v$ into a sum of elements of $D(u)$, called {\it canonical decomposition}:

\begin{equation} \label{eq:canonical} (w^tu) \cdot v=\sum_{i<j}u_{ij} c_{ij}(v, w),\text{ where $u_{ij}=e_iu_j-e_ju_i,$ $c_{ij}(v, w)=v_iw_j-v_jw_i.$}\end{equation}
Let $u, v \in R^n$ be such that $u^t v = 0$ and assume, moreover, that either $u$ or $v$ has at least one zero entry.
Recall from 3.8, 3.10 of~\cite{Ka77} that under these assumptions one can define certain element
 $x(u, v)$ of $\St(n, R)$ such that $\phi(x(u, v)) = t(u, v) = 1 + uv^t$.
The following properties of $x(u, v)$ are standard (cf.~\cite[Lemma~1.1]{Tu83}).
\begin{lemma} \label{lem:xsmall-properties}
\begin{lemlist}
 \item \label{item:xsmall-scalar} If $v$ or $w$ has at least two zero entries, then 
 \begin{equation}\nonumber x(v, wa) = x(va, w).\end{equation} %TODO: What is ''a``?
 \item \label{item:xsmall-additivity} If $w_1$ and $w_2$ have at least one common zero entry then
 \begin{equation}\nonumber x(v, w_1)x(v, w_2) = x(v, w_1+w_2).\end{equation}
 \item \label{item:xsmall-commute} If $v$, $v'$ are simultaneously orthogonal to $w$ and $w'$ and the elements $w, w'$
  both have at least two zero entries then $x(v, w)$ and $x(v', w')$ commute.
\end{lemlist}
\end{lemma}

Let $u \in E(n, R)e_1$, $v \in I^n$ be such that $u^tv = 0$.
Recall from \S~4 of~\cite{LS17} that one can define the following elements of $\St(n, R, I)$:
\begin{equation} \label{eq:sigma-definition}
 F(u, v) = \prod\limits_{i=1}^r x(u,  v_i),\ \
 S(v, u) = \prod\limits_{i=1}^r x( v_i, u).
\end{equation}
Here $\{v_r\}$, $r=1,\ldots,N$ is any collection of elements of $I^n \cap D(u)$ such that $\sum_{r=1}^N v_r = v$.
Since $u$ is unimodular, such a collection always exists by~\eqref{eq:canonical}, moreover
 $F(u, v)$ and $S(u, v)$ do not depend on its choice. %TODO: Give reference

Recall from~\cite{LS17} that it is possible to give a presentation of the relative Steinberg group $\St(n, R, I)$ 
formulated in terms of elements $F(u, v)$, $S(v, u)$.
More formally, \cite[Proposition 3.10]{LS17} asserts that $\St(n,\,R,\,I)$ admits presentation with the set of generators
\begin{multline*} \{F(u,\,v)\mid u\in E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\, \cup \\ \cup \{S(u,\,v)\mid u\in I^n,\ v\in E(n,\,R)e_1,\ u^tv=0\}\end{multline*}
subject to relations
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)^{-1}=F(t(u,\,v)u',\,t(v,\,u)^{-1} v'), \label{conj3} \\
&F(ge_1,\,g^*e_2a)=S(ge_1a,\,g^*e_2),\ \text{for all}\ a\in I,\, g \in E(n, R).
\end{align}
Here $g^*$ denotes the contragradient matrix, i.\,e. $g^* = {g^{-1}}^t$ and $e_i$ is the standard basis vector of $R^n$.

@@

To simplify notation, throughout this section we set $N_1 = N(\varpi_1)$, $N^1 = N(-\varpi_1)$.
We only construct one of the desired maps from the definition of $\sigma$-pair, namely $\sigma(\varpi_1^\vee) \colon N_1 \to N^1$,
 for shortness we denote it by $\sigma_1$.

In order to define $\sigma_1$ we will use the description of the subgroup $N_1 = N(\varpi_1^\vee)$ given in terms of ``another presentation''.
We start with the following simple observation.
\begin{lemma} \label{lem:n1-decomp} For $n\geq 4$ there is an isomorphism $N_1 \cong N_{0} \rtimes P_1^-(A)$, 
 where $N_{0}$ denotes the subgroup $\St(n, A[t], tA[t])$ and $P_1^-(A)$ is the subgroup of $\St(n, A)$ generated by $x_{ij}(\xi)$ with $i\neq 1$.
\end{lemma}

Recall that the universal property of semidirect products gives for any group $H$ acting on a group $N$ via $\phi \colon H \to \Aut(N)$ 
and any group homomorphisms $f_N\colon N \to G$, $f_H\colon H \to G$ satisfying 
\begin{equation} \label{eq:coherence-condition} f_N(\phi(h)(n)) = f_H(h) f_N(n) f_H(h)^{-1},\ (n\in N,\ h\in H) \end{equation} 
a unique map $f\colon N \rtimes H \to G$ extending $f_N$ and $f_H$.

Thus, by the above lemma, in order to construct the map $\sigma_1$ we need to construct two maps 
\[ (\sigma_1)_{P_1^-(A)} \colon P_1^-(A) \to N^1, \ \ (\sigma_1)_{N_{0}} \colon N_{0} \to N^1\]
and then verify that they satisfy~\eqref{eq:coherence-condition}.

It is easy to define the first map $(\sigma_1)_{P_1^-(A)}$, indeed, using the decomposition $P_1^-(A) \cong U^-_1 \rtimes L_1$ where 
\[U^-_1 = \langle x_{i1}(\xi) \mid i\neq 1,\ \xi\in A \rangle \text{ and } L_1 = \langle x_{ij}(\xi) \mid i,  j \neq 1,\ \xi\in A\rangle \]
we apply the universal property of semidirect products once again 
and define $(\sigma_1)_{P_1^-(A)}$ by requiring that it acts identically on $L_1$ (notice that $L_1 \subseteq N^1$) %TODO: Language???
and acts on elements of $U^-_1$ via the formula $(\sigma_1)_{P_1^-(A)}(x_{i1}(\xi))= x_{i1}(t\xi)$.

%TODO: define \delta_1 & \delta_1^{-1} -- these are not matrices

To define the map $(\sigma_1)_{N_{1,0}}$ we will invoke the presentation of the relative Steinberg group given in~\cref{sec:another-presentation}
in the special case $R=A[t]$, $I=tA[t]$.

Indeed, define $\sigma_1$ on $N_{0}=\St(n, A[t], tA[t])$ by the following identities: 
\begin{equation}
 \sigma_1(F(u, v)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot t u, \delta_1 \cdot t^{-1} v_i),\ \
 \sigma_1(S(v, u)) = \prod\limits_{i=1}^r x(\delta_1^{-1} \cdot v_i, \delta_1 \cdot u).
\end{equation}
